Constrain the Merger History of Primordial-Black-Hole Binaries from GWTC-3

### 2022-10-26 latex/paper.tex
update

\documentclass[twocolumn]{aastex631}
%\documentclass[twocolumn,linenumbers]{aastex631}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% load packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ulem}
\hypersetup{linkcolor=red,citecolor=blue,filecolor=cyan,urlcolor=magenta}
\usepackage{amsmath}
%\usepackage{subfigure}
\usepackage{float}
\usepackage{tensor}     % manipulate tensors
\usepackage{graphicx}   % include figures
\usepackage{bm}         % \bm{<text>} Bold math symbols
\usepackage{xcolor}     % textcolor
\usepackage{color}      % \textcolor{declared-color}{text}
%\usepackage{longtable}  % large table\
%\usepackage{fontspec} % accented characters for .bib file using XeLaTex
\usepackage[section]{placeins} % figures processing
\usepackage[utf8]{inputenc} % bibtex formatting

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% equations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\({\left(}
\def\){\right)}
\def\[{\left[}
\def\]{\right]}
\def\e{\begin{equation}}
\def\q{\end{equation}}
\def\m{\begin{eqnarray}}
\def\n{\end{eqnarray}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\title{The merger history of primordial-black-hole binaries inferred from GWTC-3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[0000-0001-7016-9934]{Zu-Cheng~Chen}
\email{zucheng.chen@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{You~Wu}
\email{youwuphy@gmail.com}
\affiliation{College of Mathematics and Physics, Hunan University of Arts and Science, Changde, 415000, China}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Lang~Liu}
\email{Corresponding author: liulang@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this paper, we will use the publicly available GW data from \lvc\ O1 and O2 
observations to estimate the merger rate distribution of PBH binaries with a
general mass function assuming all of \lvc\ BBHs are of primordial origin.
We find that the merger history effect makes no significant contribution to
the merger rate of PBHs and can be safely ignored.
The rest of this paper is organized as follows.
In \Sec{merger}, we review the calculation of merger rate distribution 
accounting for the merger history effect.
In \Sec{method}, we elaborate the data analysis method used to infer the
PBH populations from \lvc\ data.
In \Sec{result}, we present the results for PBHs with a power-law and 
a log-normal mass function respectively.
Finally, we summarize and discuss our results in \Sec{conclusion}.

@@

The BBHs observed by LVK suggest that BHs should have a broad mass distribution, and hence we consider an extended mass function for PBHs. Here, we denote the probability distribution function (PDF) of PBH masses by $P(m)$ and demand the PDF to be normalized such that
\e\label{norm}
    \int_{0}^{\infty} P(m)\, \rd m = 1.
\q 
Assuming the fraction of PBHs in CDM is $\fpbh$, we can estimate the abundance of PBHs in the mass interval $(m, m+\rd m)$ as \citep{Chen:2018rzo}
\e 
    0.85 \fpbh\, P(m)\, \rd m,
\q 
where the coefficient $0.85$ is roughly the fraction of CDM in non-relativistic matter including both CDM and baryons. Following \cite{Liu:2019rnx}, we may also define the ``averaged" PBH mass, $\mpbh$, as
\e\label{mpbh}
    \frac{1}{\mpbh} = \int \frac{P(m)}{m} \rd m.
\q 
Then, we can obtain the average number density of PBHs with mass $m$ in the 
total number density of PBHs, $F(m)$, by \citep{Liu:2019rnx}
\e\label{Fm} 
    F(m) = P(m) \frac{\mpbh}{m}.
\q

We can now proceed to estimate the merger rate densities of PBH binaries by considering the merger history effect.

We assume that PBHs are randomly distributed in the early Universe when they decouple from the cosmic background evolution \cite{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. Due to gravitational interactions, two nearest PBHs would attract each other and form a bound system if they get the angular momentum from the torque of the third closest PBH. Because of gravitational radiations, PBH binaries would eventually merge and may be detected by \lvc. Let us consider the first-merger process now. Assuming PBHs posses a random distribution in the early Universe, the differential distribution, for three PBHs with masses $(m_i, m_i+dm_i)$, $(m_j, m_j+dm_j)$ and  $(m_l, m_l+dm_l)$, is given by \cite{Liu:2019rnx}
\e\label{dP1}
\begin{split} 
    dP_1 =& 16\pi^2 n_T^2 x^2 y^2 F(m_i) F(m_j) F(m_l) \\
            &\times e^{-4\pi y^3 n_T/3} \rd m_i \rd m_j \rd m_l \rd x \rd y \Theta(y-x),
\end{split}
\q 
where $\Theta$ is the Heaviside step function, and $x$ is the comoving distance between two closest PBHs with masses $m_i$ and $m_j$. Note that these two nearest PBHs will form a binary due to the torque of third PBH with mass $m_l$. In \Eq{dP1}, $y$ is the comoving distance from the binary to the third PBH, and $n_T$ is the average number density of PBHs defined by
\e 
    n_T = \frac{\fpbh \rho_{\mathrm{CDM}}}{\mpbh},
\q  
where $\rho_{\mathrm{CDM}}$ is the energy density of CDM.
Notice that the $e$-factor in \Eq{dP1} guarantees that there are no other PBHs within the distance $y$. It is now straightforward to obtain the fraction of PBHs that have merged before cosmic time $t$ \cite{Liu:2019rnx}
\e 
    G_1(t) = \int \rd x \rd y \rd m_l \frac{\rd P_1}{\rd x \rd y \rd m_i \rd m_j \rd m_l} \Theta(t-\tau),
\q 
where $\tau$ is the cosmic time at which the PBH binary with masses of $m_i$ and $m_j$ merges. Therefore the merger rate density of first-merger process, $\cR_1(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \hf n_T \frac{\rd G_1(t, m_i, m_j)}{\rd t}
            = \int \hat{\cR}_1 \rd m_l,
\q 
where the factor $1/2$ accounts for the symmetry of $m_i$ and $m_j$, and
\e 
\begin{split}
    &\hat{\cR}_1(t, m_i, m_j, m_l)
    =1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
    &\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
After the first-merger process, the mass of the remnant BH will approximately be $m_i + m_j$ if the two initial PBHs have masses of $m_i$ and $m_j$.
A second-merger process would happen if this remnant BH and the nearest PBH with mass $m_k$ form a new binary by the torque from another PBH with mass $m_l$. Similarly, the merger rate density of second-merger process, $\cR_2(t, m_i, m_j)$, is given by \cite{Liu:2019rnx}
\e 
\begin{split}
    \cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
    &+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where
\e 
\begin{split}
    &\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
    =1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
    &\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
    &\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
Following the same spirit, in principle one can derive the merger rate density $\cR_n(t, m_i, m_j)$ contributed from the $n$-th merger process.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, 
of PBHs at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
    \cR(t, m_i, m_j) = \sum_{n=1} \cR_n(t, m_i, m_j),
\q 
and the total merger rate can be directly obtained by
\e\label{Rt}
    R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1} R_n(t),
\q 
where
\e 
 R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
As demonstrated in \cite{Liu:2019rnx}, $\cR_{n+1}(t, m_i, m_j)$ is not 
necessarily be smaller than $\cR_n(t, m_i, m_j)$ (see Fig.~7 and Fig.~8 in 
\cite{Liu:2019rnx}).
However, $R_{n+1}(t)$ should be smaller than $R_n(t)$ as expected
\cite{Liu:2019rnx}.
Here, we only consider the merger history up to second-merger process.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{method}Inference on PBH mass distribution from GW data}
Given a general mass function of PBHs $P(m|\vth)$ which satisfy the normalization
condition of \Eq{norm}, the time (or redshift) dependent merger rate can be
obtained by \Eq{Rt}, namely
\e\label{Rt2}
R(t|\vth) = \int \mR(t, \vla|\vth)\ \rd \vla,
\q 
where $\vla \equiv \{m_1, m_2\}$, and $\vth$ are the parameters that 
characterize the mass function and will be inferred from GW data. 
For instance, $\vth=\{\alpha, M\}$ for the power-law PDF (see \Eq{power}) 
and $\vth=\{m_c, \s\}$ for the log-normal PDF (see \Eq{log}).
The local merger rate density distribution then reads \cite{Chen:2018rzo}
\e\label{Rt0} 
\mR(t_0, \vla|\vth) = R_0\, p(\vla|\vth),
\q 
where $R_0 \equiv R(t_0|\vth)$ is the local merger rate, and $p(m_1,m_2|\vth)$ 
is the population distribution of BBH mergers.
Note that \Eq{Rt0} guarantees $p(m_1,m_2|\vth)$ is normalized, namely
\e 
    \int p(\vla|\vth)\, \rd \vla = 1.
\q

Given the GW data, $\vd = (d_1, \dots, d_N)$, which consist of $N$ BBH merger
events, we aim to extract the population parameters $\{\vth, R_0\}$ from $\vd$.
In order to do that, it is necessary to perform the hierarchical Bayesian inference
on the BBHs' mass distribution 
\cite{Abbott:2016nhf,Abbott:2016drs,TheLIGOScientific:2016pea,
    Wysocki:2018mpo,Fishbach:2018edt,Mandel:2018mve,Thrane:2018qnx}.
In this work, we will use the data of ten BBHs
\cite{TheLIGOScientific:2016pea,LIGOScientific:2018mvr} 
reported by \lvc\ O1 and O2 observations, and hence $N=10$. 
The posterior samples of these BBHs are publicly available from \footnote{\url{https://www.gw-openscience.org/}}. 
Because the standard priors on masses for each event in \lvc\ analysis are 
taken to be uniform \cite{TheLIGOScientific:2016pea,LIGOScientific:2018mvr},
the likelihood of an individual event $p(d_i|\vla)$ is
proportional to the posterior of that event $p(\vla|d_i)$.
The total likelihood for an inhomogeneous Poisson process can be evaluated as
\cite{Wysocki:2018mpo,Fishbach:2018edt,Mandel:2018mve,Thrane:2018qnx}
\e\label{likelihood}
p(\vd|\vth, R_0) \propto R_0^{N} e^{-R_0\, \beta(\vth)} \prod_i^N 
\int \rd\vla\ p(d_i|\vla)\ p(\vla|\vth),
\q 
where $\beta(\vth)$ is defined as
\e 
\beta(\vth) \equiv \int \rd\vla\ VT(\vla)\ p(\vla|\vth),
\q 
in which $VT(\vla)$ is the sensitive spacetime volume \cite{Abbott:2016nhf,Abbott:2016drs} of LIGO.
We adopt the semi-analytical approximation from 
\cite{Abbott:2016nhf,Abbott:2016drs} to estimate $VT$, 
where we use the ``IMRPhenomPv2" waveform to simulate the BBH templates
and neglect the effect of spins for BHs.
Furthermore, the threshold signal-to-noise ratio (SNR) of detection for a 
single-detector is set to $8$,
which corresponds to a network SNR threshold of around $12$.

Assuming the prior distributions $p(\vth, R_0)$ are uniform for $\vth$ parameters
and log-uniform for local merger rate $R_0$
\cite{Abbott:2016nhf,Abbott:2017vtc}, namely
\e 
p(\vth, R_0) \propto \frac{1}{R_0},
\q 
the posterior probability distribution $p(\vth, R_0|\vd)$ 
can be directly calculated by
\e\label{post} 
p(\vth, R_0|\vd) \propto p(\vd |\vth, R_0)\ p(\vth, R_0).
\q
The marginalized posterior $p(\vth|\vd)$ can then be readily obtained by
integrating over $R_0$ in \Eq{post}, namely
\e\label{post_vth} 
p(\vth|\vd) \propto \[\beta(\vth)\]^{-N} 
\prod_i^N \int \rd\vla\ p(d_i|\vla)\ p(\vla|\vth).
\q 
This marginalized posterior has been widely used in previous 
population inferences 
\cite{Abbott:2016nhf,Abbott:2017vtc,TheLIGOScientific:2016pea,
    Abbott:2016drs,Fishbach:2017zga,Chen:2018rzo}.
In the following section, we will utilize the posterior \eqref{post} to infer
the population parameters $\{\vth, R_0\}$ by considering two concrete mass
distributions, a power-law PDF and a log-normal PDF, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{result}results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Log-normal mass function}
We now consider a log-normal mass function of PBHs as \cite{Dolgov:1992pu}
\e\label{log}
P(m) = \frac{1}{\sqrt{2\pi} \s m} \exp \(-\frac{\ln^2\(m/m_c\)}{2\s^2}\),
\q
where $m_c$ presents the peak mass of $m P(m)$, and $\s$ denotes the width of 
the mass spectrum.
The log-normal mass function is a good approximation to a large class of extended mass distributions if PBHs are formed from smooth symmetric peaks in the inflationary power spectrum \cite{Carr:2017jsz}, as demonstrated in \cite{Green:2016xgy,Kannike:2017bxn}.
Note that $\vth=\{m_c, \s\}$, and the free parameters are 
$\{\vth, R_0\} = \{m_c, \s, R_0\}$ in this case. 
Using \Eq{mpbh} and \Eq{Fm}, it is easily to get
\e
\mpbh = m_c \exp \(-\frac{\s^2}{2}\),
\q
\e 
F(m) = \frac{m_c}{\sqrt{2\pi} \s m^2} \exp \(-\frac{\s^2}{2}-\frac{\ln^2\(m/m_c\)}{2\s^2}\).
\q

>>>>

\documentclass[twocolumn]{aastex631}
%\documentclass[twocolumn,linenumbers]{aastex631}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% load packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage{ulem}
%\hypersetup{linkcolor=red,citecolor=blue,filecolor=cyan,urlcolor=magenta}
\usepackage{amsmath}
%\usepackage{subfigure}
%\usepackage{float}
%\usepackage{tensor}     % manipulate tensors
%\usepackage{graphicx}   % include figures
%\usepackage{bm}         % \bm{<text>} Bold math symbols
%\usepackage{xcolor}     % textcolor
%\usepackage{color}      % \textcolor{declared-color}{text}
%\usepackage{longtable}  % large table\
%\usepackage{fontspec} % accented characters for .bib file using XeLaTex
%\usepackage[section]{placeins} % figures processing
%\usepackage[utf8]{inputenc} % bibtex formatting

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% equations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\({\left(}
\def\){\right)}
\def\[{\left[}
\def\]{\right]}
\def\e{\begin{equation}}
\def\q{\end{equation}}
\def\m{\begin{eqnarray}}
\def\n{\end{eqnarray}}
\def\permille{\ensuremath{{}^\text{o}\mkern-5mu/\mkern-3mu_\text{oo}}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% title %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\title{Constrain the Merger History of Primordial-Black-Hole Binaries from GWTC-3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[0000-0002-0297-9633]{Lang~Liu}
\email{liulang@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[0000-0002-9610-2284]{You~Wu}
\email{youwuphy@gmail.com}
\affiliation{College of Mathematics and Physics, Hunan University of Arts and Science, Changde, 415000, China}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% author %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[0000-0001-7016-9934]{Zu-Cheng~Chen}
\email{Corresponding author: zucheng.chen@bnu.edu.cn}
\affiliation{Department of Astronomy, Beijing Normal University, Beijing 100875, China}
\affiliation{Advanced Institute of Natural Sciences, Beijing Normal University, Zhuhai 519087, China}

@@

In this paper, we will use the LVK recent released GWTC-3 data to constrain the effect of merger history on the merger rate of PBH binaries assuming all of LVK BBHs are of primordial origin.
We organize the rest paper as follows.
In \Sec{merger}, we give a brief review on calculation of merger rate of PBH binaries by accounting for the merger history effect.
In \Sec{method}, we describe the hierarchical Bayesian framework used to infer the PBH population parameters from GW data.
In \Sec{result}, we consider four commonly used PBH mass functions and present the results.
Finally, we give conclusions in \Sec{conclusion}.

@@

The BBHs observed by LVK suggest that BHs should have a broad mass distribution, so we consider an extended mass function for PBHs. Here, we demand the probability distribution function of PBH mass, $P(m)$, be normalized such that
\e\label{norm}
    \int_{0}^{\infty} P(m)\, \rd m = 1.
\q 
Assuming the fraction of PBHs in CDM is $\fpbh$, we can estimate the abundance of PBHs in the mass interval $(m, m+\rd m)$ as \citep{Chen:2018rzo}
\e 
    0.85 \fpbh\, P(m)\, \rd m,
\q 
where the coefficient $0.85$ is roughly the fraction of CDM in non-relativistic matter including both CDM and baryons. Following \cite{Liu:2019rnx}, we may define an average PBH mass, $\mpbh$, as
\e\label{mpbh}
    \frac{1}{\mpbh} = \int \frac{P(m)}{m} \rd m.
\q 
Then, we can obtain the average number density of PBHs with mass $m$ in the 
total number density of PBHs, $F(m)$, by \citep{Liu:2019rnx}
\e\label{Fm} 
    F(m) = P(m) \frac{\mpbh}{m}.
\q

We can now proceed to estimate the merger rate densities of PBH binaries by considering the merger history effect.
We assume that PBHs are randomly distributed in the early Universe when they decouple from the cosmic background evolution \citep{Nakamura:1997sm,Sasaki:2016jop,Ali-Haimoud:2017rtz}. Two nearest PBHs would attract each other because of the gravitational interactions. These two PBHs would obtain the angular momentum from the torque of other PBHs and form a PBH binary. The binary would emit gravitational radiations and eventually merge.

We do not intend to give a detailed derivation, but quote the results from \cite{Liu:2019rnx} here.
The merger rate density from first-merger process, $\cR_1(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\cR_1(t, m_i, m_j) = \int \hat{\cR}_1 \rd m_l,
\q 
where
\e 
\begin{split}
    &\hat{\cR}_1(t, m_i, m_j, m_l)
    \equiv 1.32 \times 10^6 \times \(\frac{t}{t_0}\)^{-\frac{34}{37}}\(\frac{\fpbh}{\mpbh}\)^\frac{53}{37} \\ 
    &\times m_l^{-\frac{21}{37}} (m_i m_j)^\frac{3}{37} (m_i + m_j)^\frac{36}{37} F(m_i) F(m_j) F(m_l).     
\end{split}
\q
Similarly, the merger rate density from second-merger process, $\cR_2(t, m_i, m_j)$, is given by \citep{Liu:2019rnx}
\e 
\begin{split}
    \cR_2(t, m_i, m_j) &= \hf \int \hat{\cR}_2(t, m_i-m_e, m_e, m_j, m_l)\ \rd m_l \rd m_e\\
    &+ \hf \int \hat{\cR}_2(t, m_j-m_e, m_e, m_i, m_l)\ \rd m_l \rd m_e,
\end{split}
\q 
where
\e 
\begin{split}
    &\hat{\cR}_2(t, m_i, m_j,m_k, m_l)
    =1.59 \times 10^4 \times \(\frac{t}{t_0}\)^{-\frac{31}{37}} \(\frac{\fpbh}{\mpbh}\)^\frac{69}{37}\\ 
    &\qquad\times m_k^{\frac{6}{37}} m_l^{-\frac{42}{37}} (m_i+ m_j)^\frac{6}{37} (m_i + m_j + m_k)^\frac{72}{37}\\
    &\qquad\times F(m_i) F(m_j) F(m_k) F(m_l).     
\end{split}
\q
We only consider the effect of merger history up to second-merger process.
Therefore the total merger rate density, $\cR(t, m_i, m_j)$, of PBH binaries at cosmic time $t$ with masses $m_i$ and $m_j$ is
\e\label{cR}
    \cR(t, m_i, m_j) =  \sum_{n=1,2} \cR_n(t, m_i, m_j),
\q 
and the total merger rate is
\e\label{Rt}
    R(t) = \int \cR(t, m_i, m_j) \rd m_i \rd m_j = \sum_{n=1,2} R_n(t),
\q 
where
\e 
 R_n(t) = \int \cR_n(t, m_i, m_j) \rd m_i \rd m_j.
\q 
All the above mentioned merger rate (density) is measured at source frame.
We should emphasize that although $R_2(t)$ should be smaller than $R_1(t)$ as expected, $\cR_2(t, m_i, m_j)$ is not necessarily be smaller than $\cR_1(t, m_i, m_j)$ \citep{Liu:2019rnx}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{method}Hierarchical Bayesian Inference}
To infer the population parameters, we adopt a hierarchical Bayesian approach by marginalizing over the uncertainty in the estimate of individual event parameters.
In this section, we describe the hierarchical Bayesian inference used in the parameter estimations.
The merger rate density \eqref{cR} is measured in source frame, and we need to convert it into the detector frame as
\e\label{Rpop}
\mR_{\mathrm{pop}}(\theta|\Ld) = \frac{1}{1+z} \frac{dV_\mathrm{c}}{dz} \mR(\theta|\Ld),
\q
where $z$ is the cosmological redshift, $\theta\equiv\{z, m_1, m_2\}$, $\Ld$ is a collection of $\fpbh$ and the parameters from mass function $P(m)$, and $dV_\mathrm{c}/dz$ is the differential comoving volume. The factor $1/(1 + z)$ converts time increments from the source frame to the detector frame. In this work, we take the cosmological parameters from Planck 2018 \citep{Planck:2018vyg}.

Given the data, $\textbf{d} = \{d_1, d_2, \cdots, d_{N_{\mathrm{obs}}}\}$, of $N_{\mathrm{obs}}$ BBH merger events, we model the total number of events as an inhomogeneous Poisson process, yielding the likelihood \citep{Loredo:2004nn,Thrane:2018qnx,Mandel:2018mve}
\begin{equation}\label{L1}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{\int \mathcal{L} (d_{i}| \theta)\, \cR_{\mathrm{pop}}(\theta|\Ld) d \theta}{\xi(\Ld)},
\end{equation}
where $N_{\exp}\equiv N_{\exp}(\Ld)$ is the expected number of detections over the timespan of observation, and $\mathcal{L} (d_{i}|\theta)$ is the individual event likelihood for the $i$th GW event that can be derived from the individual event's posterior by reweighing with the prior on $\theta$. Here, $\xi(\Ld)$ quantifies selection biases for a population with parameters $\Ld$ and is defined by
\e\label{xi}
\xi(\Ld) = \int P_{\mathrm{det}}(\theta)\, \cR_{\mathrm{pop}}(\theta|\Ld)\, \mathrm{d} \theta,
\q
where $P_{\text{det}}(\theta)$ is the detection probability that depends on the source parameters $\theta$.
In practice, we use the simulated injections \citep{ligo_scientific_collaboration_and_virgo_2021_5546676} to estimate $\xi(\Ld)$, and \Eq{xi} can be approximated by a Monte Carlo integral over found injections \citep{LIGOScientific:2021psn}
\begin{equation}
	\xi(\Ld) \approx \frac{1}{N_{\mathrm{inj}}} \sum_{j=1}^{N_{\text {found }}} \frac{\cR_{\mathrm{pop}}(\theta_{j} | \Ld)}{p_{\mathrm{draw}}(\theta_j)},
\end{equation}
where $N_{\text{inj}}$ is the total number of injections, $N_{\text{found}}$ is the number of successfully detected injections, and $p_{\mathrm{draw}}$ is the probability density function from which the injections are drawn.
Using the posterior samples from each individual event, the hyper-likelihood \eqref{L1} can be estimated as
\begin{equation}\label{L2}
	\mathcal{L}(\textbf{d}|\Ld) \propto N_{\mathrm{exp}}^{N_{\mathrm{obs}}} e^{-N_{\mathrm{exp}}} \prod_{i=1}^{N_{\mathrm{obs}}} \frac{1}{\xi(\Ld)} \left\langle \frac{\cR_{\mathrm{pop}}(\theta|\Ld)}{d_L^2(z)} \right\rangle,
\end{equation}
where $\langle\cdots\rangle$ denotes the weighted average over posterior samples of $\theta$. The denominator $d_L^2(z)$ is the standard priors used in the LVK analysis of individual events where $d_L$ is the luminosity distance.

In this work, we incorporate the PBH population distribution \eqref{cR} into the \texttt{ICAROGW} \citep{Mastrogiovanni:2021wsd} package to estimate the likelihood function \eqref{L2}, and use \texttt{dynesty} \citep{Speagle:2019ivv} sampler called from \texttt{Bilby} \citep{Ashton:2018jfp,Romero-Shaw:2020owr} to sample over the parameter space. We use the GW events from GWTC-3 by discarding events with false alarm rate larger than 1 yr$^{-1}$ and events with the secondary component mass smaller than $3\Msun$ to avoid contamination from putative events involving neutron stars following \cite{DeLuca:2021wjr}. A total of $69$ GW events from GWTC-3 meet this criteria and the posterior samples of these BBHs are publicly available from \cite{ligo_scientific_collaboration_and_virgo_2021_5655785}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\label{result}results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[tbp!]
	\centering
	\begin{tabular}{lll}
		\hline\hline
		\textbf{Parameter\quad} & \textbf{Description} & \textbf{Prior} \\
		\hline
		$\fpbh$ & Abundance of PBH in dark matter & log-$\mU(-4, 0)$\\
		\hline
		\multicolumn{3}{c}{Lognormal PBH mass function} \\[1pt]
		$\Mc$ & Central mass in $\Msun$. & $\mU(5, 50)$\\
		$\sigma$ & Mass width. & $\mU(0.1, 2)$\\
		\hline
		\multicolumn{3}{c}{Power-law PBH mass function} \\[1pt]
		$\Mmin$ & Lower mass cut-off in $\Msun$. & $\mU(3, 10)$\\
		$\al$ & Power-law index. & $\mU(1.05, 4)$\\
		\hline
		\multicolumn{3}{c}{Broken Power-law PBH mass function} \\[1pt]
		$M_*$ & Peak mass in $\Msun$. & $\mU(3, 15)$\\
		$\al_1$ & First power-law index. & $\mU(0, 15)$\\
		$\al_2$ & Second power-law index. & $\mU(1, 3)$\\
		\hline
		\multicolumn{3}{c}{Critical collapse (CC) PBH mass function} \\[1pt]
		$\Mf$ & Horizon mass scale in $\Msun$. & $\mU(1, 50)$\\
		$\al$ & Universal exponent. & $\mU(0, 5)$\\
		\hline
	\end{tabular}	
	\caption{\label{table:priors}Parameters and their prior distributions used in the Bayesian parameter estimations. Here, $\mU$ and log-$\mU$ denote uniform and log-uniform distributions, respectively.}
\end{table}
Based on the hierarchical statistical framework, we do the parameter estimations for four different PBH mass functions that are commonly used in the literature. These mass functions are the log-normal, power-law, broken power-law, and critical collapse (CC) distributions, respectively. We summarize the parameters and their prior distributions in \Table{table:priors}. Below we show the results for each of the PBH mass function.

\subsection{Log-normal mass function}
We first consider a PBH mass function with the log-normal form \citep{Dolgov:1992pu} of
\e\label{log}
P(m) = \frac{1}{\sqrt{2\pi} \s m} \exp \(-\frac{\ln^2\(m/M_c\)}{2\s^2}\),
\q
where $M_c$ represents the central mass of $m P(m)$, and $\s$ characterizes the width of the mass spectrum.
A huge class of extended mass distributions can be approximated by the log-normal mass function if PBHs are formed from a smooth symmetric peak in the inflationary power spectrum when the slow-roll approximation holds \citep{Green:2016xgy,Carr:2017jsz,Kannike:2017bxn}.
The hyper-parameters are $\Ld = \{M_c, \s, \fpbh\}$ in this case. 
We can then derive the averaged PBH mass and averaged number density from \Eq{mpbh} and \Eq{Fm} as
\e
\mpbh = M_c \exp \(-\frac{\s^2}{2}\),
\q
\e 
F(m) = \frac{M_c}{\sqrt{2\pi} \s m^2} \exp \(-\frac{\s^2}{2}-\frac{\ln^2\(m/M_c\)}{2\s^2}\).
\q

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=0.5\textwidth]{post-log.pdf}
	\caption{\label{posterior-log}The marginalized one- and two-dimensional posterior distributions for hyper-parameters $\{M_c, \s, \fpbh\}$ in the log-normal mass function inferred from GWTC-3. The blue color denotes the results from the first merger only, while the orange denotes the results from both the first and second mergers. The contours represent the $1\sigma$, $2\sigma$ and $3\sigma$ credible regions, respectively.}
\end{figure}

\begin{figure}[htbp!]
	\centering
	\includegraphics[width=0.5\textwidth]{ratio-log.pdf}
	\caption{\label{ratio-log}The ratio of merger rate density from second merger to that from first merger, $\cR_2(t_0, m_1, m_2)/\cR_1(t_0, m_1, m_2)$, as a function of component masses for the log-normal mass function. We have fixed the hyper-parameters $\{M_c, \s, \fpbh\}$ to their best-fit values.}
\end{figure}

Using $69$ BBHs from GWTC-3 and performing the hierarchical Bayesian inference, we obtain $M_c = 17.3^{+2.2}_{-2.0} \Msun$, $\s = 0.71^{+0.10}_{-0.08}$, and $\fpbh = 1.8^{+0.3}_{-0.3} \times 10^{-3}$. In this work, we present result with median value and 90\% equal-tailed credible intervals. The posteriors for the hyper-parameters $\Ld = \{M_c, \s, \fpbh\}$ are shown in \Fig{posterior-log}. Note that we get a larger value of $M_c$ than that inferred from GWTC-1 in \cite{Wu:2020drm} because GWTC-3 contains more heavier BHs than those from GWTC-1. From \Eq{Rt}, we also infer the local merger rate as $R(t_0) = 41^{+16}_{-12} \gpcyr$. The results of local merger rate and abundance of PBHs are consistent with the previous estimations \citep{Chen:2018czv,Chen:2018rzo,Chen:2019irf,Wu:2020drm,Chen:2021nxo,Chen:2022fda}, confirming that CDM cannot be dominated by the stellar-mass PBHs.
